Prediction of Drinking Water Treatment Process Chemical Doses

### 2025-12-17 Paper Draft.Rmd
Incorporated Josh's GAM. Wrote conclusion from bullet points that I presented in slide. Re-formatted table in Ethan's section. Added descriptions to all classification sections and some of the dose prediction sections. Added description to exploratory analysis section.

```{r, include=F}
# Import required libraries
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
library(ggplot2)
library(zoo)
library(purrr)
library(knitr)
library(patchwork)
library(tidyverse)
library(tidymodels)
library(rpart.plot)
library(kableExtra)
library(vip)
library(png)
library(grid)
library(parsnip)
library(discrim)
library(gt)
library(Metrics)
```

@@

## Project Purpose
***The goal of this project is to produce a model that serves the role of a 
treatment plant operator.*** This model would provide recommended chemical doses,
based on input water quality characteristics, and how operators have dosed 
chemicals in the past. This could prove a useful tool to operators as a "starting
point" in chemical dosing. They could see a change in water quality, retrieve
the suggested chemical dose from the model, test this dose, and adjust from there
based on other information.

@@

* Raw Water Data
    + pH
    + Temperature (of the water)
    + Turbidity
    + Suspended Grain Size Information
    + Alkalinity
    + Hardness

@@

* **pH**: No significant outliers that we could see, so we simply day-averaged.
* **Temperature**: There was one region of approximately zero temperature (unrealistic
for the reservoir). We got rid of these values and interpolated to the nearest non-zero temperature.
* **Conductivity**: Values less than 0.2 and greater than 2 were removed, since they 
are unreasonable based on this dataset and the "usual" values.
* **Turbidity**: A few exceptionally high values (relative to the usual data values)
were removed before averaging.
* **Grain Size Information**: This data was highly variable, and simply looks like "noise",
so did not seem helpful for prediction. We thus did not clean it.
* **Alkalinity and Hardness**: These data have different time availability than the other
datasets but will be inspected. Removed zero-values.

@@

Some of this cleaned data is presented below.

@@

Alkalinity and hardness tend to stay fairly steady over time.

@@

@@

@@

Based on preliminary analyses, we will focus on the following dose predictors:

@@

Outcome: **FeCl Dose**. (Polymer dose is not changed significantly in plants.)

@@

We introduced **lagged variables** like in the Stock dataset from the homework.

To capture time-dependence of this dataset, we will introduce lagged variables. 
For each of the main water parameters, we will introduce two lagged variables, giving
the water characteristics for the day before and two days before.

We also introduced a **categorical predictor** indicating whether the dose increased, decreased, or stayed the same compared to the day before.

@@

### Training and Testing Datasets
- Training and Testing
   - Training 2018-2019, Testing 2020
   - Approximately 70/30 split

### Models Used
The following models will be used for the two main tasks at hand. Models that work well with numerical outcomes will be used for coagulant dose prediction. Models that are effective for classification will be used for coagulant dose change prediction, since this is a three-category classification problem.

- **Coagulant Dose Prediction**
   - Linear Regression Models
   - Tree-Based Models
   - Generalized Additive Model
- **Coagulant Dose Change Prediction**
   - LDA and QDA Classification
   - Multinomial Classification
   - Random Forest Classification

@@

@@

@@

```{r, out.width="80%", out.height="88%"}
# Creating Table and Top Model Graphs

@@

kable_table <- performance %>%
  kable("html", caption = "Model Performance Summary") %>%
  kable_styling(full_width = F, position = "center") %>%
  row_spec(which(performance$RMSE == best_rmse), background = "#cfe2f3")

kable_table

@@

p3
p4
```

@@

```{r, fig.width=5, fig.height=2.5}
set.seed(445)

@@

Random forest was used was the logical next step to try to better capture the nonlinear patterns in our data. Multiple different tree amounts were attempted, but ultimately they all performed similarly. Although the random forest was unable to model the data, we were able to determine which features had the most importance. A graph was created to visualize the feature importance for this dataset. Notably, pH and the pH lag variables were the most important. Temperature and the temperature lag variables followed, which is similar to what we concluded from other models and our exploratory analysis.

@@

**Josh's section to insert his GAM**

@@

```{r, fig.height=3.8, fig.width=7}

@@

dosepred_df <- data.frame(Model = c("Random Forest", "Decision Tree", "Cond-pH + Cond-Temp", "Conductivity-Temp", "Temp and pH"),
                          RMSE = c(2.811, 2.756, 1.672, 1.891, 2.350),
                          MAE = c(2.283, 2.484, 1.421, 1.516, 1.941))

@@

@@

The goal of dose change prediction models is to predict whether coagulant dose increased, decreased, or stayed the same based on raw water characteristics.

@@

The following is a helpful visualization of how dose change varies over time with each of the key raw water characteristics.

```{r, fig.width=7, fig.height=3}

@@

Based on this plot, [in essence, we're screwed, but say it diplomatically]

### LDA

- Experimented with predictor sets and found that pH and temperature (plus the associated lagged variables) were the main effective predictors.
- Turbidity and conductivity did not improve prediction.

LDA ACCURACY: 43.7%

@@

# performance metrics
lda_fit_disp <- lda_fit |> extract_fit_engine()
lda_conf <- lda_fit |> 
  augment(new_data = df_test) |>
  conf_mat(truth = FeCl_change, estimate = .pred_class)
lda_acc <- lda_fit |>
  augment(new_data = df_test) |>
  accuracy(truth = FeCl_change, estimate = .pred_class)

@@

### QDA

- Experimented with predictor sets and found that pH and temperature (plus the associated lagged variables) were the main effective predictors.
- Turbidity and conductivity did not improve prediction.

QDA ACCURACY: 35.7%

@@

# performance metrics
qda_fit_disp <- qda_fit |> extract_fit_engine()
qda_conf <- qda_fit |> 
  augment(new_data = df_test) |>
  conf_mat(truth = FeCl_change, estimate = .pred_class)
qda_acc <- qda_fit |>
  augment(new_data = df_test) |>
  accuracy(truth = FeCl_change, estimate = .pred_class)

@@

### Multinomial Classification

- Experimented with predictor sets and found that pH and temperature (plus the associated lagged variables) were the main effective predictors.
- Turbidity and conductivity did not improve prediction.

MULTINOMIAL ACCURACY: 46.3%

@@

multi_recipe1 <- recipe(FeCl_change ~ mean_Temp + Temp_lag1 + Temp_lag2 + mean_pH + pH_lag1 + pH_lag2 + mean_Conductivity + Cond_lag1 + Cond_lag2, data = df_train) |>
  step_normalize(all_numeric_predictors()) |>
  step_dummy(all_nominal_predictors())

@@

## NOTES
# turbidity is highly variable and does not help

@@

- Experimented with predictor sets and found that pH and temperature (plus the associated lagged variables) were the main effective predictors.
- Turbidity and conductivity did not improve prediction.

RANDOM FOREST ACCURACY: 39.2%

@@

## results
rf_acc <- rf_fit |>
  augment(new_data = df_test) |>
  accuracy(truth = FeCl_change, estimate = .pred_class)
rf_conf <- rf_fit |>
  augment(new_data = df_test) |>
  conf_mat(truth = FeCl_change, estimate = .pred_class)

@@

## Notes
# 34% accuracy for all predictors
# 39% accuracy for temp + pH + temp/pH lags
# 46.6% accuracy for just temperature and pH

@@

Overall ineffective, and prone to overfitting (especially random forest).  
Model predicting no dose change would be more accurate.

```{r, fig.width=7, fig.height=3}

@@

ggplot(dosechangepred_df, aes(x = reorder(Model,-Accuracy), y = Accuracy)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = scales::percent(Accuracy, accuracy = 0.1)),
            vjust = -0.3, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, max(dosechangepred_df$Accuracy) * 1.1)) +
  labs(x = "Model", y = "Accuracy") +
  theme_minimal()

@@

# Overview of Model Results

(talk about how the models compared, more just synthesizing than making commentary. save commentary for conclusion)

@@

- Predicting chemical doses based on these raw water characteristics proved to be ineffective
- Most likely reasons:
   - Limited size of dataset
   - Limited predictor set
   - Operator-dependent
   - Complexity of chemical relationships
   - Other factors (example: reservoir turnover)
- Operators make decisions based on variety of other factors: chemical waste production, filter performance, lab tests, etc.
- Incorporate more varieties and time ranges of data in future modeling efforts

>>>>

```{r, include=F}
# Import required libraries
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
library(ggplot2)
library(zoo)
library(purrr)
library(knitr)
library(patchwork)
library(tidyverse)
library(tidymodels)
library(rpart.plot)
library(kableExtra)
library(vip)
library(png)
library(grid)
library(parsnip)
library(discrim)
library(gt)
library(Metrics)
library(mgcv)
```

@@

## Project Purpose
***The goal of this project is to produce a model that serves the role of a 
treatment plant operator.*** This model would provide recommended chemical doses,
based on input water quality characteristics, and how operators have dosed 
chemicals in the past. This could prove a useful tool to operators as a "starting
point" in chemical dosing. They could see a change in water quality, retrieve
the suggested chemical dose from the model, test this dose, and adjust from there
based on other information. We will attempt to create both a predictive model for predicting the chemical dose, and a model to predict dose change (increase, decrease, stay same), both based on raw water characteristics.

@@

* Raw Water Data
    + pH
    + Temperature
    + Turbidity ("cloudiness" of the water)
    + Suspended Grain Size Distributions
    + Alkalinity (resistance of water to changes in pH)
    + Hardness (mineral content of water)

@@

* **pH**: No significant outliers were present that we could see, so we simply day-averaged.
* **Temperature**: There was one region of approximately zero temperature (unrealistic
for the reservoir). We got rid of these values and interpolated to the nearest non-zero temperature.
* **Conductivity**: Values less than 0.2 and greater than 2 were removed, since they 
are unreasonable based on this dataset and the "usual" values.
* **Turbidity**: A few exceptionally high values (relative to the usual data values)
were removed before averaging.
* **Grain Size Information**: This data was highly variable, and simply looks like "noise",
so did not seem helpful for prediction. We thus did not clean it.
* **Alkalinity and Hardness**: These data have different time availability than the other
datasets but will be inspected. Removed zero-values.

@@

@@

While alkalinity and hardness display seasonal trends that could be helpful for prediction, the available time ranges of the data did not align with other datasets (namely, dosing and other raw water characteristics), so they will not be used in this analysis. However, they are pictured below. Note the gap in data availability from mid-2017 to mid-2018.

@@

The main four raw water characteristics used in this analysis are pictured below. Turbidity displays the highest levels of fluctuation and noise, followed by conductivity. pH and temperature tend to display fairly consistent annual cycles without significant noise. According to the treatment operator who provided the data, temperature and pH are important for the treatment process, so less noise could be advantageous for modeling.

@@

The following are the dosing over time of Ferric Chloride (primary coagulant) and the polymer coagulant. Note the long periods of no dosing change, some periods of sharp dips and jumps, and periods of steady increase and decrease. This is how operators alter coagulant doses day-to-day. This gives a picture of some of the challenges that may be present in trying to predict this coagulant dosage based on water characteristics.

@@

Based on preliminary analyses, we will focus on the following dose predictors, which have the greatest time availability and fewest data gaps out of our available raw water data:

@@

We will focus on **FeCl Dose** as the outcome variable for our modeling efforts. Polymer dose is not changed significantly in plants, as it has less direct impact on the coagulation performance and serves as a secondary "helper" in particle formation to the ferric chloride.

@@

To capture some of the time-dependence likely present in our dataset, we introduced **lagged variables** for each of our main raw water characteristics. We introduced two lagged variables, representing the water characteristics for the day before and two days before.

We also introduced a **categorical predictor** indicating whether the dose increased, decreased, or stayed the same compared to the day before. This will be used for dose change prediction modeling (classification), which is detailed below.

@@

### Training and Testing Datasets
Our total dataset spans 2018-2020, so we will use data from 2018-2019 as the training dataset, and 2020 as the testing dataset. This accomplishes approximately a 70/30 training/testing split for our available data.

We recognize that there may be limitations to this train/test split, since 2020 was both the height of the COVID-19 pandemic, and a year of above-average numbers of forest fires. The former of these could have impacted the funding and staffing of the water treatment plant, along with consumptive use patterns in the regions serviced by the plant. The fires could have directly impacted water quality with more dissolved ash in the water supply.

These could have unforeseen impacts, but it seems upon preliminary assessment that these factors should not impact plant performance significantly. Water treatment is critical infrastructure, so should not be substantially impacted by shutdowns. Additionally, it is already only staffed by a few people at a time. Shifting consumptive use patterns would not impact reservoir water quality.

Reservoirs also tend to be fairly stable water sources, in terms of quality, so the additional ash that may have come from the fires is likely negligible. We observed some sparse datasets available, including dissolved organic carbon and total organic carbon measurements, and 2020 was not significantly different from 2018 or 2019 in this way, which could be a helpful indicator.

With this in mind, this split will be maintained. For future efforts, with more data, different splits will be tested, but with this available data, this seems like the best approach.

### Models Used
The following models will be used for the two main tasks at hand. Models that work well with numerical outcomes will be used for coagulant dose prediction. Models that are effective for classification will be used for coagulant dose change prediction, since this is a three-category classification problem (increase, decrease, stay same).

- **Coagulant Dose Prediction**
   - Linear Regression Models
   - Tree-Based Models
   - Generalized Additive Model
- **Coagulant Dose Change Prediction**
   - LDA and QDA Classification
   - Multinomial Logistic Regression
   - Random Forest Classification

Test MAE and RMSE will be used to assess dose prediction models, with best model chosen based on RMSE. Test accuracy will be used to assess dose change prediction models.

@@

**The goal of the dose prediction modeling is to predict FeCl dosage from the raw water characteristics.** The following plot helps illustrate the data that will be used for the dose prediction model. This plot does not reveal any clear trends in dosage relative to any of the water treatment variables (dose varying with a water quality metric), indicating this could be a challenging modeling effort.

@@

Displayed are both a summary of the fitted regression models, and an example plot of the actual versus predicted dosing, presented for the best-performing linear model, Model 3. This better illustrates how far the predictions are from the actual dosing.

@@

## Create graphs and get model results

@@

kable_table_pdf <- performance %>%
  # Use format = "latex" for PDF output
  kable("latex",
        booktabs = TRUE)

@@

kable_table_pdf

p3

@@

A decision tree was fit with a depth of 5, and 50 observations required to split. These parameters were varied but it did not significantly improve prediction. The model split first based on pH, which was mentioned by treatment plant operators as an important predictor. However, the next splits were on lagged conductivity and turbidity measurements. This was surprising, and could likely be due to the noise present in the data. (Possible result of overfitting.) This model performed more poorly than the linear models. See below for model performance metrics and a visualization of the final decision tree.

```{r, fig.width=8, fig.height=4}
set.seed(445)

@@

Random forest was used was the logical next step to try to better capture the nonlinear patterns in our data. Multiple different tree amounts were attempted, but ultimately they all performed similarly, and worse than the linear models. Although the random forest was unable to model the data, we were able to determine which features had the most importance. A graph was created to visualize the feature importance for this dataset. Notably, pH and the pH lag variables were the most important. Temperature and the temperature lag variables followed, which is similar to what we concluded from other models and our exploratory analysis.

@@

Next, a GAM was fitted to the modeled data. This model uses splines to model nonlinear relationships between predictor and response. The hope would be that this flexible model form could help capture the nonlinearity present in these chemical processes, for effective prediction. See below for a plot of how the GAM performed on the test dataset.

train <- df_lag_incdec |> filter(year(Timestamp) <= 2019)
test  <- df_lag_incdec |> filter(year(Timestamp) == 2020)

gam_spec <- gen_additive_mod() |> 
  set_engine("mgcv") |> 
  set_mode("regression")
gam_rec <- recipe(FeCl_dose ~ Polymer_dose + mean_Temp + mean_pH + mean_Conductivity +
                 mean_Turbidity + Temp_lag1 + Temp_lag2 +
                 pH_lag1 + pH_lag2 + Cond_lag1 + Cond_lag2 +
                 Turb_lag1 + Turb_lag2, data = train)

gam_wf <- workflow() |>
  add_model(gam_spec) |>
  add_recipe(gam_rec)

gam_fit <- gam(
  FeCl_dose ~ s(Polymer_dose) + s(mean_Temp) + s(mean_pH) + s(mean_Conductivity) +
               s(mean_Turbidity) + s(Temp_lag1) + s(Temp_lag2) +
               s(pH_lag1) + s(pH_lag2) + s(Cond_lag1) + s(Cond_lag2) +
               s(Turb_lag1) + s(Turb_lag2),
  data = train
)

# gam_fit |> tidy()
# gam_fit |> glance()

pred_df <- test %>% 
  mutate(pred = predict(gam_fit, newdata = test))

gam_rmse <- yardstick::rmse(pred_df, truth = FeCl_dose, estimate = pred)
gam_mae <- yardstick::mae(pred_df, truth = FeCl_dose, estimate = pred)

ggplot(pred_df, aes(x = Timestamp)) +
  geom_line(aes(y = FeCl_dose, color = "Actual")) + 
  geom_line(aes(y = pred, color = "Predicted")) + 
  scale_color_manual(
    name = "FeCl Dose",
    values = c("Actual" = "orange", "Predicted" = "skyblue")) +
  labs(
    y = "FeCl Dose (mg/L)" # Add a proper Y-axis label
  ) +
  theme_minimal() +
  ggtitle("FeCl Dosing Predictions by GAM on Test Dataset")

Overall, the GAM was able to vaguely follow trends in dosing for the testing dataset, but did not perform overly well, with MAE and RMSE of 1.8 mg/L and 2.3 mg/L, respectively. The model flexibility did not turn out to yield the hoped-for results, though it performed similarly to the standard linear models. It did out-perform both of the tree-based models.

@@

Overall, none of the dose prediction models performed particularly well or were able to follow the changes in chemical dose present in the modeling. The top three models by RMSE were the models involving interactions between conductivity and temperature/pH, and the GAM. However, the best-performing models still differed from the actual dosing by an average of at least 1.5 mg/L based on MAE, which is a significant dosing deviation that could hinder proper functioning of a treatment plant if implemented.

```{r, fig.height=4.2, fig.width=7}

@@

dosepred_df <- data.frame(Model = c("Random Forest", "Decision Tree", "Cond-pH + Cond-Temp", "Conductivity-Temp", "Temp and pH", "GAM"),
                          RMSE = c(2.811, 2.756, 1.672, 1.891, 2.350, gam_rmse$.estimate),
                          MAE = c(2.283, 2.484, 1.421, 1.516, 1.941, gam_mae$.estimate))

@@

Next, we will move on to dose change prediction, and see how its performance compares to this model estimation of exact doses.

@@

**The goal of dose change prediction models is to predict whether coagulant dose increased, decreased, or stayed the same based on raw water characteristics.**

@@

The following is a helpful visualization of how dose change varies over time with each of the key raw water characteristics. As with dose prediction, there are not clear patterns of when dose is increasing, decreasing, or staying the same relative to water quality metrics. This will most likely make the data difficult to separate via classification methods.

@@

For each of the following classification models, different combinations of predictor variable sets were experimented with. It was found that pH and temperature, with their associated lag variables, produced the highest test accuracies. Turbidity and conductivity did not improve test accuracy, so were not included in the final model that is represented by the confusion matrices and accuracies below.

### Linear Discriminant Analysis (LDA)

LDA uses linear combinations of features to perform dimensionality reduction and create better separation between classes of data. Thus, we hope that for this hard-to-separate data, the dimension reduction could assist, though there are not many dimensions to start with, so we could foresee it still being highly limited.

TEST ACCURACY: 43.7%

@@

# performance metrics
lda_fit_disp <- lda_fit |> extract_fit_engine()
lda_conf <- lda_fit |> 
  augment(new_data = df_test) |>
  conf_mat(truth = FeCl_change, estimate = .pred_class)
lda_acc <- lda_fit |>
  augment(new_data = df_test) |>
  yardstick::accuracy(truth = FeCl_change, estimate = .pred_class)

@@

### Quadratic Discriminant Analysis (QDA)

QDA is similar to LDA but incorporates quadratic terms, not just linear terms. This could potentially help address the nonlinearity in our dataset.

However, based on test accuracy and the confusion matrix, the QDA model performed just as poorly as the LDA, and slightly worse. See results below.

TEST ACCURACY: 35.7%

@@

# performance metrics
qda_fit_disp <- qda_fit |> extract_fit_engine()
qda_conf <- qda_fit |> 
  augment(new_data = df_test) |>
  conf_mat(truth = FeCl_change, estimate = .pred_class)
qda_acc <- qda_fit |>
  augment(new_data = df_test) |>
  yardstick::accuracy(truth = FeCl_change, estimate = .pred_class)

@@

### Multinomial Logistic Regression

Multinomial logistic regression is an extension of logistic regression that allows for multi-category classification, by setting one category as a "reference category" and calculating probabilities of other categories relative to that category. Overall, this model tended to predict the best on the test dataset. Looking at the confusion matrix, it tended to lean towards predicting "same" when it made an error, which is more acceptable than choosing the opposite direction of dose change, though it made that error as well. However, it is still not an exceptional model by any means. See below for confusion matrix and accuracy.

TEST ACCURACY: 46.0%

@@

multi_recipe1 <- recipe(FeCl_change ~ mean_Temp + Temp_lag1 + Temp_lag2 + mean_pH + pH_lag1 + pH_lag2, data = df_train) |>
  step_normalize(all_numeric_predictors()) |>
  step_dummy(all_nominal_predictors())

@@

@@

Finally, a Random Forest model was fit to this data. The standard number $\sqrt{P}$ predictors were sampled at each split, and 500 trees were grown. The number of trees was varied but it did not change prediction accuracy.

This model performed second-worst out of the classification models. We believe that this is mainly due to overfitting. Random forest models tend to overfit on noise-heavy data, so this model performed perfectly on the training dataset and very poorly on the test dataset. See accuracy and confusion matrix, below.

TEST ACCURACY: 39.2%

@@

## results
rf_acc <- rf_fit |>
  augment(new_data = df_test) |>
  yardstick::accuracy(truth = FeCl_change, estimate = .pred_class)
rf_conf <- rf_fit |>
  augment(new_data = df_test) |>
  conf_mat(truth = FeCl_change, estimate = .pred_class)

@@

@@

These models were ineffective at predicting the direction of dose change. This aligns with initial expectations, since the data was shown to be extremely scattered and difficult to separate in the time series plots shown at the start of this section. All models performed very poorly, so it is hard to compare which models may have performed "best." At the surface, multinomial performed best, likely due to it using linear combinations of predictors (which aligns with the most effective dose prediction methods). Random forest was the most prone to overfitting and worst-performing, which aligns with its tendency to overfit on noise-heavy data. Model accuracies are summarized below.

```{r, fig.width=7, fig.height=3.2}

@@

ggplot(dosechangepred_df, aes(x = reorder(Model,-Accuracy), y = Accuracy)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = scales::percent(Accuracy, accuracy = 0.1)),
            vjust = -0.3, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, max(dosechangepred_df$Accuracy) * 1.1)) +
  labs(x = "Model", y = "Accuracy") +
  theme_minimal() +
  ggtitle("Dose Change Classification Model Comparison")

@@

@@

Overall, predicting chemical doses based on these raw water characteristics proved to be ineffective.

The dose prediction models produced high RMSE's, either being unable to capture the complexity of the dosing, or overfitting to the noise-filled signals of the input data. The dose change prediction models, as predicted by observation of the timeseries plots, yielded extremely low accuracies, due to this being a dataset not separable by any one of these four input water quality characteristics.

Some of the most likely reasons for this are as follows:

* **Limited dataset size**: our dataset was limited in time, forcing us to use a small training and testing dataset that limited our ability to choose other years for testing. This likely produced errors specific to the year that we chose for testing, which could be averted with more data.  
* **Limited predictor set**: These four water quality variables were not effective for predicting chemical doses. However, there are other water quality parameters, and other treatment plant performance characteristics such as filter performance, that are used to make dose decisions. Access to these data could significantly improve model performance.  
* **Operator-dependence**: Whether doses even change at all is largely dependent on operators. This plant is one of the more advanced in the nation, but some plants leave their dose the same for years at a time. So, our model is in part simulating human behavior, which is much more difficult than a basic model can capture.  
* **Complexity of chemical relationships**: The drinking water treatment process is an enormous chemical reaction with a large number of interconnected processes and chemical interactions, so our models likely failed in part to not being able to fully capture these relationships.  
* **Other factors**: Other miscellaneous factors can impact water quality, such as reservoir turnover, which is when a reservoir's water level drops, and water quality changes drastically over the course of a few days at a specific time of year. This is hard to predict, and our model could not capture this.

Ultimately, water treatment is both a science and an art, and operators make decisions based on variety of inputs: chemical waste production, filter performance, or lab tests, to name a few. Future modeling efforts should incorporate more varieties and time ranges of data to attempt to capture some of these complexities.
